Hardware-Accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-06-23 paper.tex
Consolidated paragraphs about slicing method variants (ref #17)
First pass at redoing evaluation prose (ref #17)

As a baseline the Boolean-based approaches were compared to the existing
view-aligned slicing method from previous work by developing a simple prototype.

In addition, a variation of the slicing method that was optimized for two
volumes was created.  Basically rather than using uniform variables to switch
textures as would normally be done, the optimized variation stores the texture
with each vertex instead.  While it requires a fairly significant amount of
extra storage, using vertex attributes allows all the slices to be transferred
to the card at once.  The result is a considerable performance boost, similar to
how it is generally faster to transfer one large file across a network then many
small ones.

As shown both Boolean-based approaches outperformed the traditional slicing
method, but were in turn bested by the optimized slicing method.  It should be
noted that the frame rates for the optimized slicing method are the worst case,
and on average it performed even better, around forty frames per second.

Judging from the numbers, it appears that the multiple passes of the
Boolean-based approaches create too much overhead compared to the optimized
slicing method.  The optimized slicing method also has a very simple fragment
shader which may keep it from being fragment-bound.

@@

Altogether, although the Boolean-based approaches outperform the un-optimized
slicing method, it is recommended that the traditional slicing method with the
optimizations described be used.  The optimized slicing method is not only
faster, but it supports volumes of any orientation.  Furthermore, the optimized
method should scale to more volumes relatively easily.

However, it is hoped that the lessons learned in this paper, namely that the
overhead incurred by multi-pass algorithms may outweigh the cost of streaming
geometry in other algorithms, will prove valuable to future endeavors in both
volume rendering and graphics in general.

>>>>

As a baseline the Boolean-based approaches were compared to the existing
view-aligned slicing method from previous work by developing a simple prototype.
To be fair, a variation of the slicing method optimized for two volumes was
created.  Basically rather than using uniform variables to switch textures as
would normally be done, the optimized variation stores the texture with each
vertex instead.  While it requires a fairly significant amount of extra storage,
using vertex attributes allows all the slices to be transferred to the card at
once.  The result is a considerable performance boost, similar to how it is
generally faster to transfer one large file across a network then many small
ones.

As shown the Boolean-based approaches generally outperformed the two variants of
the slicing method.  The difference between the frame rates becomes more clear
as the number of samples increases.

Judging from the numbers, it appears all the approaches are fragment-bound, and
thus the Boolean-based approaches win out because they can employ early ray
termination.

@@

Altogether, it is difficult to say for sure which approach developers should
take to render intersecting volumes.  If the application only needs to handle
two axis-aligned volumes, using one of the Boolean-based approaches could
provide a more interactive experience for the user.  However, if the application
needs more than two volumes, or if those volumes need to be oriented
arbitrarily, it is best to stick with the slicing method.

In the future it would be interesting to see if either of the two Boolean-based
approaches could be expanded to support more volumes and different orientations.
The Boolean XOR approach would most likely be easier to extend to more volumes,
since it is just a matter of creating more pieces.  However, the depth-masking
approach will be better off for supporting different orientations because it
works in pixel space.

Either way, it is hoped that the lessons learned in this report, namely that
rendering volumes is inherently fragment-bound and avoiding any unnecessary
processing with early ray termination is key to achieving significant speed-ups,
will prove valuable to future future endeavors in both volume rendering and
graphics in general.
